Using Deep Learning to Identify Patients with Cognitive Impairment in Electronic Health Records

### 2021-11-24 Submission/Paper/ML4H/Paper.tex

% use the "wcp" class option for workshop and conference
 % proceedings
 %\documentclass[gray]{jmlr} % test grayscale version
 %\documentclass[tablecaption=bottom]{jmlr}% journal article
 \documentclass[pmlr,twocolumn,10pt]{jmlr} % W&CP article

@@

\author{Tanish Tyagi\nametag{\thanks{Authors contributed equally}\SUP{1}},
Colin G. Magdamo\nametag{\footnotemark[1]\SUP{1}}, 
Ayush Noori\SUP{1},
Zhaozhi Li\SUP{1},
Xiao Liu\SUP{1},
Mayuresh Deodhar\SUP{1},
Zhuoqiao Hong\SUP{1},
Wendong Ge\SUP{1},
Elissa M. Ye\SUP{1},
Yi-han Sheu\SUP{1},
Haitham Alabsi\SUP{1},
Laura Brenner\SUP{1},
Gregory K. Robbins\SUP{1},
Sahar Zafar\SUP{1}, 
Nicole Benson\SUP{1},
Lidia Moura\SUP{1},
John Hsu\SUP{1},
Alberto Serrano-Pozo\SUP{1},
Dimitry Prokopenko\SUP{1},
Rudolph E. Tanzi\SUP{1},
Bradley T. Hyman\SUP{1},
Deborah Blacker\SUP{1},
Shibani S. Mukerji\SUP{1},
M. Brandon Westover\SUP{1},
Sudeshna Das\SUP{1}
\centering \Email{
\\[\bigskipamount] 
\SUP{1}\{ttyagi, 
cmagdamo,
anoori1,
zli39,
xliu61,
mdeodhar,
zhong1,
wendong.ge,
emye,
ysheu,
halabsi,
lnbrenner,
grobbins,
sfzafar, 
nbenson, 
lidia.moura, 
john.hsu,
aserrano1,
dprokopenko,
dblacker,
rtanzi,
bhyman,
smukerji,
mwestover,
sdas5\}
@mgh.harvard.edu}
\centering
\begin{center}\addr Massachusetts General Hospital, Boston, MA\end{center}
}

@@

\paragraph{Dataset}
\label{sec:Dataset} Our dataset consisted of a cohort (N = 16,428) of patients from the Mass General Brigham (MGB) HealthCare \citep{jpm6010011} (formerly Partner's Healthcare, comprising two major academic hospitals, community hospitals, and community health centers in the Boston area) system who were older than 60 years (as of July 13, 2021), had \textit{APOE} genotypes—the biggest genetic risk factor for Alzheimer's disease  \citep{mahley2000apolipoprotein}—available from the MGB BioBank, and at least one clinician note with a dementia-related keyword. \tableref{tab:table1} shows demographics of the cohort of patients.

@@

\label{sec:Transformer}  
\paragraph{(2) Transformer Based Sequence Classification Language Model} We utilized a pre-trained language model called ClinicalBERT \citep{alsentzer2019publicly}, which was trained on the MIMIC II \citep{saeed2011multiparameter} database containing EHR records from ICU patients. We used the implementation in the Huggingface Transformers \citep{Wolf2019HuggingFacesTS} and Simpletransformers \citep{simple2020thilina} packages. After text preprocessing, input texts were tokenized with the default tokenizer and converted to embeddings. The model was initialized with pre-trained parameters and later fine-tuned on our labeled training set. Optuna \citep{akiba2019optuna} was used to perform a 20-trial study and tune the learning rate, Adam epilson, and the number of train epochs on the held-out validation set to maximize AUC. An early stopping rule was used to prevent overfitting by ensuring that training stopped if the loss did not change substantially over 3 epochs.

@@

\begin{figure}[htb]
\caption{Confusion Matrix Patient Level Prediction Counts for ClinicalBERT}
\resizebox{\columnwidth}{!}{%
\label{fig:fig2}
\centering 
\includegraphics[scale=0.5]{bert_patient_predictions.png}

>>>>

% use the "wcp" class option for workshop and conference
 % proceedings
 %\documentclass[gray]{jmlr} % test grayscale version
 %\documentclass[tablecaption=bottom]{jmlr}% journal article
 %\pdfoutput=1
 \documentclass[pmlr,twocolumn,10pt]{jmlr} % W&CP article

@@

\author{Tanish Tyagi\nametag{\thanks{Authors contributed equally}\SUP{1}},
Colin G. Magdamo\nametag{\footnotemark[1]\SUP{1}}, 
Ayush Noori\SUP{1},
Zhaozhi Li\SUP{1},
Xiao Liu\SUP{1},
Mayuresh Deodhar\SUP{1},
Zhuoqiao Hong\SUP{1},
Wendong Ge\SUP{1},
Elissa M. Ye\SUP{1},
Yi-han Sheu\SUP{1},
Haitham Alabsi\SUP{1},
Laura Brenner\SUP{1},
Gregory K. Robbins\SUP{1},
Sahar Zafar\SUP{1}, 
Nicole Benson\SUP{1},
Lidia Moura\SUP{1},
John Hsu\SUP{1},
Alberto Serrano-Pozo\SUP{1},
Dimitry Prokopenko\SUP{1},
Rudolph E. Tanzi\SUP{1},
Bradley T. Hyman\SUP{1},
Deborah Blacker\SUP{1},
Shibani S. Mukerji\SUP{1},
M. Brandon Westover\SUP{1},
Sudeshna Das\SUP{1}
\centering \Email{
\\[\bigskipamount] 
\SUP{1}\{ttyagi, 
cmagdamo,
anoori1,
zli39,
xliu61,
mdeodhar,
zhong1,
wendong.ge,
emye,
ysheu,
halabsi,
lnbrenner,
grobbins,
sfzafar, 
nbenson, 
lidia.moura, 
john.hsu,
aserrano1,
dprokopenko,
dblacker,
rtanzi,
bhyman,
smukerji,
mwestover,
sdas5\}
@mgh.harvard.edu}
\centering
\begin{center}\addr Massachusetts General Hospital and McCance Center for Brain Health, Boston, MA 
\end{center}
}

@@

\paragraph{Dataset}
\label{sec:Dataset} Our dataset consisted of a cohort (N = 16,428) of patients from the Mass General Brigham (MGB) HealthCare  (formerly Partner's Healthcare, comprising two major academic hospitals, community hospitals, and community health centers in the Boston area) system who were older than 60 years (as of July 13, 2021), had \textit{APOE} genotypes—the biggest genetic risk factor for Alzheimer's disease  \citep{mahley2000apolipoprotein}—available from the MGB BioBank, and at least one clinician note with a dementia-related keyword. \tableref{tab:table1} shows demographics of the cohort of patients.

@@

\label{sec:Transformer}  
\paragraph{(2) Transformer Based Sequence Classification Language Model} We utilized a pre-trained language model called ClinicalBERT \citep{alsentzer-etal-2019-publicly}, which was trained on the MIMIC II \citep{saeed2011multiparameter} database containing EHR records from ICU patients. We used the implementation in the Huggingface Transformers \citep{Wolf2019HuggingFacesTS} and Simpletransformers \citep{simple2020thilina} packages. After text preprocessing, input texts were tokenized with the default tokenizer and converted to embeddings. The model was initialized with pre-trained parameters and later fine-tuned on our labeled training set. Optuna \citep{akiba2019optuna} was used to perform a 20-trial study and tune the learning rate, Adam epilson, and the number of train epochs on the held-out validation set to maximize AUC. An early stopping rule was used to prevent overfitting by ensuring that training stopped if the loss did not change substantially over 3 epochs.

@@

\begin{figure}[htb]
\caption{Confusion Matrix Patient Level Prediction Counts for ClinicalBERT}
\resizebox{\columnwidth}{!}{%
\label{fig:fig2}
\centering 
\includegraphics[scale=0.5]{Confusion_Matrix.png}
